---
title: "macro factors 1 with MSRP and Inflation"
author: "Vamshi Gadepally"
date: "2023-06-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Import Libraries}
library(fpp)
library(TSA)
library(tseries)
library(ggplot2)
library(forecast)
library(caret)
library(lubridate)
library(dplyr)
library(AICcmodavg)
library(vars)
```

```{r Q1}
# loading the data
data <- read.csv("C:\\Users\\vamsh\\Documents\\Uni\\UChicago\\Summer 2023\\MSCA 34003_IP01 - Capstone II\\Project\\Data\\Macro Factors\\merged_table_final_filtered_no_Iron_Ore.csv")
```


#########
### Time series analysis on Macro factor data
#########


```{r Q1}
# # Check for missing values
# missing_values <- sapply(data, function(x) sum(is.na(x)))
# missing_values_data <- data.frame(Variable = names(data), Missing_Values = missing_values)
# missing_values_data # No missing values
```

```{r Q1}
# tsdisplay(data$Average_MSRP_spec)
# # tsdisplay(data$CCI)
# # tsdisplay(data$GDP)
# # tsdisplay(data$Metal)
# # tsdisplay(data$Iron_Ore)
# # tsdisplay(data$Inflation)
```

```{r Q6}
# Creating a copy for EDA
data_2<-data.frame(data)
```

```{r}
# # Setting Covid_period to 1 for months between and including 2020-01 to 2020-12
# data_2$Covid_period[format(data_2$DATE) >= "2020-03" &
#                      format(data_2$DATE) <= "2020-06"] <- 1
# 
# # Setting Covid_period to 0 for all other months
# data_2$Covid_period[format(data_2$DATE) < "2020-03" |
#                       format(data_2$DATE) > "2020-06"] <- 0
```

```{r Q1}
# Splitting the data into training and testing sets
# Note: this it the data that wasn't detrended / deseasonalized or transformed

# Train/Test Split
data_2$DATE <- as.Date(paste0(data_2$DATE, "-01"))
data_3 <- subset(data_2, select = c("DATE", "Average_MSRP_spec", "Inflation"))

# # Create the train dataframe (Jan 2013 to June 2020) - 90 months
# train_start <- as.Date("2013-01-01")
# train_end <- as.Date("2020-06-01")
# train_df <- subset(data_3, DATE >= train_start & DATE <= train_end)
# 
# # Create the test dataframe (July 2020 to June 2021) - 12 months
# test_start <- as.Date("2020-07-01")
# test_end <- as.Date("2021-06-01")
# test_df <- subset(data_3, DATE >= test_start & DATE <= test_end)

# Create the train dataframe (Jan 2013 to Apr 2021) - 100 months
train_start <- as.Date("2013-01-01")
train_end <- as.Date("2021-04-01")
train_df <- subset(data_3, DATE >= train_start & DATE <= train_end)

# Create the test dataframe (May 2021 to Apr 2022) - 12 months
test_start <- as.Date("2021-05-01")
test_end <- as.Date("2022-04-01")
test_df <- subset(data_3, DATE >= test_start & DATE <= test_end)
```

//////////////////////////////////////////////////////////////////////////////////////////////////////////////////////////////////////////////////////////////

```{r Q1}
# converting split data sets to time series objects
train_ts <- ts(train_df$Average_MSRP_spec)
test_ts <- ts(test_df$Average_MSRP_spec)
tsdisplay(train_ts)
```

```{r Q1}
# # Augmented Dickey-Fuller Test
# adf.test(train_ts) # Appears to be stationary
```

The p-value (<0.01) is less than the significance level (0.05) for Average_MSRP_spec
Which means that the null hypothesis can be rejected, and the time series is considered stationary.

```{r Q1}
# # KPSS test
# kpss.test(train_ts, null = c("Level", "Trend"), lshort = TRUE) # Appears to be stationary
```

The p-value is 0.1.
Since this value is not less than .05, we fail to reject the null hypothesis of the KPSS test.
This means we can assume that the time series is stationary.

```{r Q1}
# # plotting the time series data
# tsdisplay(train_ts, main = 'Average_MSRP_spec (training set) Jan 2013 - Sep 2019', xlab = 'Time') 
# # there is no clear seasonality present in the Average_MSRP_spec data
```

```{r Q1}
# lambda_value = BoxCox.lambda(train_ts)
# lambda_value # -0.9999242
```

```{r Q13}
# # Plotting the Box-Cox Transformed time series with the best lambda
# train_ts_trans <- BoxCox(train_ts, lambda_value)
# tsdisplay(train_ts_trans, main = 'Box-Cox Transformation Average_MSRP_spec with ð€= -0.9999242', xlab = 'Time') 
```

The BoxCox.lambda() function proposed an appropriate lambda value for the data, but it's close to 1 (no transformation is 
required), suggesting the original data didn't have significant changes (increase or decrease) in variation with the level of
the time series.

```{r Q14}
# # 1st Differencing (ð’… = ðŸ)
# tsdisplay(diff(train_ts_trans), main = '1st Differencing of Average_MSRP_spec', xlab = 'Time') 
# # 1st order Differencing isn't required
```

//////////////////////////////////////////////////////////////////////////////////////////////////////////////////////////////////////////////////////////////

```{r}
#multivariate linear regression with all variables
Average_MSRP_spec <- train_ts

Inflation <- train_df[,"Inflation"]

# Fit a linear model
lin_model <- tslm(Average_MSRP_spec ~ Inflation)
summary(lin_model)
pacf(lin_model$residuals)
```

```{r Q1}
# SARIMA/ARIMA with multiple regression errors
Average_MSRP_spec <- train_ts
exogenous <- train_df[, c("Inflation")]
# Convert the exogenous variables to a matrix
exogenous_matrix <- as.matrix(exogenous)

# Fit the auto.arima model with xreg
ARIMAX_model<- auto.arima(Average_MSRP_spec, xreg = exogenous_matrix) 
summary(ARIMAX_model)
pacf(ARIMAX_model$residuals)
```

```{r Q5}
# Vector Autogressive Model
train_ts_2 <- ts(train_df[, -1])
test_ts_2 <- ts(test_df[, -1])
# Determine the optimal lag order using the 'VARselect' function
max_lag <- 10  # Maximum lag order to consider
var_select <- VARselect(train_ts_2, lag.max = max_lag, type = "both")
# Get the optimal lag order based on AIC
lag_order <- var_select$selection[1]
# Fit the VAR model
var_model <- VAR(train_ts_2, p = lag_order, type = "both")
summary(var_model)
# Note: checkresiduals() function is only for univariate time series models
```

```{r Q5}
#Preparing xreg variables
xreg_vars <- test_df[, c("Inflation")]

#Forecasting
fit_forecast_ARIMAX <- forecast(ARIMAX_model, xreg= xreg_vars ,h=length(test_ts))
fit_forecast_lin <- forecast(lin_model, h=length(test_ts), newdata = data.frame(test_df))
fit_forecast_VAR <- forecast(var_model, h=length(test_ts), newdata = data.frame(test_df[, -1]))
```

```{r Q27}
data_forecast_ARIMAX <- data.frame(Time = test_df$DATE,
                   Actual = test_df$Average_MSRP_spec,
                   Forecasted = fit_forecast_ARIMAX)

data_forecast_lin <- data.frame(Time = test_df$DATE,
                   Actual = test_df$Average_MSRP_spec,
                   Forecasted = fit_forecast_lin)

data_forecast_VAR <- data.frame(Time = test_df$DATE,
                   Actual = test_df$Average_MSRP_spec,
                   Forecasted = fit_forecast_VAR$forecast$Average_MSRP_spec)
```

```{r Q27}
# Plot the forecasted vs. actual values for Linear model
ggplot(data_forecast_lin, aes(Time)) +
  geom_ribbon(aes(ymin = Forecasted.Lo.95, ymax = Forecasted.Hi.95),
              fill = "#BFD4E9", alpha = 0.5) +
  geom_ribbon(aes(ymin = Forecasted.Lo.80, ymax = Forecasted.Hi.80),
              fill = "#9FC5E8", alpha = 0.5) +
  geom_line(aes(y = Actual, color = "Actual")) +
  geom_line(aes(y = Forecasted.Point.Forecast, color = "Forecasted")) +
  labs(x = "Time", y = "Average_MSRP_spec") +
  scale_color_manual(values = c(Actual = "red", Forecasted = "blue")) +
  theme_minimal() + ggtitle("12 Month Forecast vs. Actual Values for Linear model with Inflation")
```

```{r Q27}
# Accuracy for 12 Month - Linear model
forecast::accuracy(fit_forecast_lin, test_df$Average_MSRP_spec) %>% round(4)
```

```{r Q27}
# residual plot for Linear model
checkresiduals(lin_model)
```

The Ljung-Box tests for the presence of residual autocorrelation. The test resulted in a p-value of greater than 0.05, 
indicating no evidence of autocorrelation. 
Also when we examine the residual plot, we observe no significant spikes in the Autocorrelation Function (ACF) at any lags. 
Spikes indicate residual patterns that are not accounted for by the model. Model adequately captures patterns in the data.

```{r Q27}
# Plot the forecasted vs. actual values for ARIMAX model
ggplot(data_forecast_ARIMAX, aes(Time)) +
  geom_ribbon(aes(ymin = Forecasted.Lo.95, ymax = Forecasted.Hi.95),
              fill = "#BFD4E9", alpha = 0.5) +
  geom_ribbon(aes(ymin = Forecasted.Lo.80, ymax = Forecasted.Hi.80),
              fill = "#9FC5E8", alpha = 0.5) +
  geom_line(aes(y = Actual, color = "Actual")) +
  geom_line(aes(y = Forecasted.Point.Forecast, color = "Forecasted")) +
  labs(x = "Time", y = "Average_MSRP_spec") +
  scale_color_manual(values = c(Actual = "red", Forecasted = "blue")) +
  theme_minimal() + ggtitle("12 Month Forecast vs. Actual Values for ARIMAX model with Inflation")
```

```{r Q27}
# Accuracy for 12 Month - ARIMAX model
forecast::accuracy(fit_forecast_ARIMAX, test_df$Average_MSRP_spec) %>% round(4)
```

```{r Q27}
# residual plot for ARIMAX model
checkresiduals(ARIMAX_model)
```

The Ljung-Box test resulted in a p-value of greater than 0.05 for this model too, indicating no evidence of autocorrelation. 
Also when we examine the residual plot, we observe no significant spikes in the Autocorrelation Function (ACF) at any lags. 
Model adequately captures patterns in the data.

```{r Q27}
# Plot the forecasted vs. actual values for VAR model
ggplot(data_forecast_VAR, aes(Time)) +
  geom_ribbon(aes(ymin = Forecasted.Lo.95, ymax = Forecasted.Hi.95),
              fill = "#BFD4E9", alpha = 0.5) +
  geom_ribbon(aes(ymin = Forecasted.Lo.80, ymax = Forecasted.Hi.80),
              fill = "#9FC5E8", alpha = 0.5) +
  geom_line(aes(y = Actual, color = "Actual")) +
  geom_line(aes(y = Forecasted.Point.Forecast, color = "Forecasted")) +
  labs(x = "Time", y = "Average_MSRP_spec") +
  scale_color_manual(values = c(Actual = "red", Forecasted = "blue")) +
  theme_minimal() + ggtitle("12 Month Forecast vs. Actual Values for VAR model with Inflation")
```

```{r Q27}
# Accuracy for 12 Month - Linear model
forecast::accuracy(fit_forecast_VAR$forecast$Average_MSRP_spec, test_df$Average_MSRP_spec) %>% round(4)
```

```{r}
# # Saving model and loading for later use
# saveRDS(lin_model, "./lin_model_Time_Series.rds")
# # lin_model <- readRDS("lin_model_Time_Series.rds")
# 
# saveRDS(ARIMAX_model, "./ARIMAX_model_Time_Series.rds")
# # ARIMAX_model <- readRDS("ARIMAX_model_Time_Series.rds")
# 
# saveRDS(var_model, "./var_model_Time_Series.rds")
# # var_model <- readRDS("lin_model_Time_Series.rds")
```


#########
### Regression analysis on Macro factor data
#########


```{r Q27}
# Create the random seed number for reproducible results
seedNum <- 888
control <- trainControl(method="repeatedcv", number=10, repeats=1)
```

### Generating models using linear algorithms
```{r LR}
# Linear Regression (Regression)
set.seed(seedNum)
# Run algorithms using 10-fold cross validation
fit.lm <- train(Average_MSRP_spec~., data=train_df[, -1], method="lm", metric="RMSE", trControl=control)
print(fit.lm)
summary(fit.lm)
checkresiduals(fit.lm)
```

```{r RIDGE}
# # Ridge (Regression)
# set.seed(seedNum)
# fit.ridge <- train(Average_MSRP_spec~., data=train_df[, -1], method="ridge", metric="RMSE", trControl=control)
# print(fit.ridge)
# summary(fit.ridge)
# checkresiduals(fit.ridge)
```

```{r LASSO}
# # The Lasso (Regression)
# set.seed(seedNum)
# fit.lasso <- train(Average_MSRP_spec~., data=train_df[, -1], method="lasso", metric="RMSE", trControl=control)
# print(fit.lasso)
# summary(fit.lasso)
# checkresiduals(fit.lasso)
```

```{r ENET}
# # ElasticNet Regression
# set.seed(seedNum)
# fit.enet <- train(Average_MSRP_spec~., data=train_df["Inflation"], method="enet", metric="RMSE", trControl=control)
# print(fit.enet)
# summary(fit.enet)
# checkresiduals(fit.enet)
```

### Generating models using nonlinear algorithms

```{r CART}
# Decision Tree - CART (Regression/Classification)
set.seed(seedNum)
fit.cart <- train(Average_MSRP_spec~., data=train_df[, -1], method="rpart", metric="RMSE", trControl=control)
print(fit.cart)
summary(fit.cart)
checkresiduals(fit.cart)
```

### Generating models using ensemble algorithms
We can tune these algorithms to see whether we can improve the results.

```{r BAGCART}
# Bagged CART (Regression/Classification)
set.seed(seedNum)
fit.bagcart <- train(Average_MSRP_spec~., data=train_df[, -1], method="treebag", metric="RMSE", trControl=control)
print(fit.bagcart)
summary(fit.bagcart)
checkresiduals(fit.bagcart)
```

```{r RF}
# Random Forest (Regression/Classification)
set.seed(seedNum)
fit.rf <- train(Average_MSRP_spec~., data=train_df[, -1], method="rf", metric="RMSE", trControl=control)
print(fit.rf)
summary(fit.rf)
checkresiduals(fit.rf)
```

```{r GBM}
# Gradient Boosting (Regression/Classification)
set.seed(seedNum)
fit.gbm <- train(Average_MSRP_spec~., data=train_df[, -1], method="xgbTree", metric="RMSE", trControl=control, verbose=F, verbosity = 0)
# fit.gbm <- train(Average_MSRP_spec~., data=train_df[, -1], method="gbm", metric="RMSE", trControl=control, verbose=F)
print(fit.gbm)
summary(fit.gbm)
checkresiduals(fit.gbm)
```

### Improving Accuracy or Results
The two best-perfoming algorithms were the Lasso and ElasticNet Regression models. We will now Search for a combination of 
parameters for each algorithm that yields the best results.

```{r FINAL1}
# # Tuning algorithm #1 - Lasso model
# set.seed(seedNum)
# parameters <- c(seq(0.1, 2, by =0.1) ,  seq(2, 5, 0.5) , seq(5, 25, 1))
# grid <- expand.grid(alpha=1, lambda = parameters)
# fit.lasso2 <- train(Average_MSRP_spec~., data=train_df[, -1], method="glmnet", metric="RMSE", tuneGrid=grid, trControl=control)
# plot(fit.lasso2)
# print(fit.lasso2)
# summary(fit.lasso2)
# checkresiduals(fit.lasso2)
```

```{r FINAL1}
# # Tuning algorithm #2 - ElasticNet model
# set.seed(seedNum)
# fractionTune = seq(.3,.8,.1)
# lambdaTune = seq(.05, .3, .1)
# grid <- expand.grid(lambda = lambdaTune, fraction = fractionTune)
# fit.enet2 <- train(Average_MSRP_spec~., data=train_df[, -1], method="enet", metric="RMSE", tuneGrid=grid, trControl=control)
# plot(fit.enet2)
# print(fit.enet2)
# summary(fit.enet2)
# checkresiduals(fit.enet2)
```

```{r CART}
# # Get model coefficients
# predict(fit.enet2$finalModel, type = c("fit", "coefficients"))
```

```{r FINAL1}
# Tuning algorithm #4 - Decision Tree - CART model
set.seed(seedNum)
# Define the hyperparameter grid using expand.grid()
grid <- expand.grid(cp = seq(0.01, 0.3, by = 0.01))   # Complexity parameter (cost complexity pruning parameter)
fit.cart2 <- train(Average_MSRP_spec~., data=train_df[, -1], method="rpart", metric="RMSE", tuneGrid=grid, trControl=control)
plot(fit.cart2)
print(fit.cart2)
summary(fit.cart2)
checkresiduals(fit.cart2)
```

```{r FINAL1}
# Tuning algorithm #5 - Bagged CART model
set.seed(seedNum)
# Perform hyperparameter tuning
fit.bagcart2 <- train(Average_MSRP_spec ~ ., data = train_df[, -1], method = "treebag",  metric = "RMSE", trControl = control, nbagg = 106)
#plot(fit.bagcart2)
print(fit.bagcart2)
summary(fit.bagcart2)
checkresiduals(fit.bagcart2)
```

```{r FINAL1}
# Tuning algorithm #6 - Random Forest model
set.seed(seedNum)
# Define the hyperparameter grid
hyperparameters <- expand.grid(
  mtry = (1:15)) # Number of variables randomly sampled as candidates at each split

# Perform hyperparameter tuning
fit.rf2 <- train(Average_MSRP_spec ~ ., data = train_df[, -1], method = "rf", metric = "RMSE", trControl = control, 
                tuneGrid = hyperparameters)
plot(fit.rf2)
print(fit.rf2)
summary(fit.rf2)
checkresiduals(fit.rf2)
```

```{r FINAL1}
# Tuning algorithm #7 - Gradient Boosting model
set.seed(seedNum)
# Define the hyperparameter grid
hyperparameters <- expand.grid(max_depth = c(2, 3, 4, 5, 6, 7),                # Maximum tree depth
                        nrounds = (1:10)*50,       # Number of boosting iterations
                        # default values below
                        eta = 0.3,                 # Learning rate
                        gamma = 0,                 # Minimum loss reduction required for a split
                        subsample = 1,             # Subsample ratio of the training instances
                        min_child_weight = 1,      # Minimum sum of instance weight needed in a child
                        colsample_bytree = 0.6)    # Subsample ratio of columns when constructing each tree

# Perform hyperparameter tuning
fit.gbm2 <- train(Average_MSRP_spec ~ ., data = train_df[, -1], method = "xgbTree", metric = "RMSE", trControl = control, verbose = FALSE, verbosity = 0, tuneGrid = hyperparameters)
plot(fit.gbm2)
print(fit.gbm2)
summary(fit.gbm2)
checkresiduals(fit.gbm2)
```

### Comparing algorithms
```{r SPOT_CHECK}
# # RIDGE=fit.ridge, LASSO=fit.lasso, ENET=fit.enet, 
# results <- resamples(list(LR=fit.lm, CART=fit.cart, BagCART=fit.bagcart, RF=fit.rf, GBM=fit.gbm))
# summary(results)
# dotplot(results)
# cat('The average RMSE from all models is:',
#     mean(c(results$values$`LR~RMSE`,results$values$`RIDGE~RMSE`,results$values$`LASSO~RMSE`,results$values$`ENET~RMSE`,results$values$`CART~RMSE`,results$values$`BagCART~RMSE`,results$values$`RF~RMSE`,results$values$`GBM~RMSE`)),'\n')
```

### Compare Algorithms Before and After Tuning

```{r POST_TUNING}
# Ridge=fit.ridge, Lasso=fit.lasso2, ElasticNet=fit.enet2, Ridge_1=fit.ridge, Lasso_1=fit.lasso, ElasticNet_1=fit.enet had errors
results_post_tuning <- resamples(list(CART_2=fit.cart2, Bagged_CART_2=fit.bagcart2, Random_Forest_2=fit.rf2, Gradient_Boosting_2=fit.gbm2, LR=fit.lm, CART_1=fit.cart, Bagged_CART_1=fit.bagcart, Random_Forest_1=fit.rf, Gradient_Boosting_1=fit.gbm))
summary(results_post_tuning)
dotplot(results_post_tuning)
```

```{r Q5}
# # Get the number of observations used for training
# n <- nrow(train_df)
# 
# # Make point predictions using the trained model on the test data
# point_predictions <- predict(fit.ridge, newdata = test_df)
# 
# # Get the estimated standard error of the residuals
# residuals <- test_df$Average_MSRP_spec - point_predictions
# sigma_hat <- sqrt(sum(residuals^2) / (n - length(fit.ridge$finalModel$lambda)))
# 
# # Calculate the confidence intervals for predictions
# confidence_level_80 <- 0.8
# confidence_level_95 <- 0.95
# 
# z_value_80 <- qnorm(1 - (1 - confidence_level_80) / 2)
# z_value_95 <- qnorm(1 - (1 - confidence_level_95) / 2)
# 
# confidence_intervals_80 <- data.frame(
#   lower_80 = point_predictions - z_value_80 * sigma_hat,
#   upper_80 = point_predictions + z_value_80 * sigma_hat
# )
# 
# confidence_intervals_95 <- data.frame(
#   lower_95 = point_predictions - z_value_95 * sigma_hat,
#   upper_95 = point_predictions + z_value_95 * sigma_hat
# )
# 
# # Combine point predictions and confidence intervals into a single dataframe
# pred_CI_ridge <- cbind(point_predictions, confidence_intervals_80, confidence_intervals_95)
```

```{r Q5}
# # Get the number of observations used for training
# n <- nrow(train_df)
# 
# # Make point predictions using the trained model on the test data
# point_predictions <- predict(fit.lasso2, newdata = test_df)
# 
# # Get the estimated standard error of the residuals
# residuals <- test_df$Average_MSRP_spec - point_predictions
# sigma_hat <- sqrt(sum(residuals^2) / (n - length(fit.lasso2$finalModel$lambda)))
# 
# # Calculate the confidence intervals for predictions
# confidence_level_80 <- 0.8
# confidence_level_95 <- 0.95
# 
# z_value_80 <- qnorm(1 - (1 - confidence_level_80) / 2)
# z_value_95 <- qnorm(1 - (1 - confidence_level_95) / 2)
# 
# confidence_intervals_80 <- data.frame(
#   lower_80 = point_predictions - z_value_80 * sigma_hat,
#   upper_80 = point_predictions + z_value_80 * sigma_hat
# )
# 
# confidence_intervals_95 <- data.frame(
#   lower_95 = point_predictions - z_value_95 * sigma_hat,
#   upper_95 = point_predictions + z_value_95 * sigma_hat
# )
# 
# # Combine point predictions and confidence intervals into a single dataframe
# pred_CI_lasso2 <- cbind(point_predictions, confidence_intervals_80, confidence_intervals_95)
```

```{r Q5}
# # Get the number of observations used for training
# n <- nrow(train_df)
# 
# # Make point predictions using the trained model on the test data
# point_predictions <- predict(fit.enet2, newdata = test_df)
# 
# # Get the estimated standard error of the residuals
# residuals <- test_df$Average_MSRP_spec - point_predictions
# sigma_hat <- sqrt(sum(residuals^2) / (n - length(fit.enet2$finalModel$lambda)))
# 
# # Calculate the confidence intervals for predictions
# confidence_level_80 <- 0.8
# confidence_level_95 <- 0.95
# 
# z_value_80 <- qnorm(1 - (1 - confidence_level_80) / 2)
# z_value_95 <- qnorm(1 - (1 - confidence_level_95) / 2)
# 
# confidence_intervals_80 <- data.frame(
#   lower_80 = point_predictions - z_value_80 * sigma_hat,
#   upper_80 = point_predictions + z_value_80 * sigma_hat
# )
# 
# confidence_intervals_95 <- data.frame(
#   lower_95 = point_predictions - z_value_95 * sigma_hat,
#   upper_95 = point_predictions + z_value_95 * sigma_hat
# )
# 
# # Combine point predictions and confidence intervals into a single dataframe
# pred_CI_enet2 <- cbind(point_predictions, confidence_intervals_80, confidence_intervals_95)
```


```{r Q5}
# Get the number of observations used for training
n <- nrow(train_df)

# Make point predictions using the trained model on the test data
point_predictions <- predict(fit.cart2, newdata = test_df)

# Get the estimated standard error of the residuals
residuals <- test_df$Average_MSRP_spec - point_predictions
sigma_hat <- sqrt(sum(residuals^2) / (n - length(fit.cart2$finalModel$lambda)))

# Calculate the confidence intervals for predictions
confidence_level_80 <- 0.8
confidence_level_95 <- 0.95

z_value_80 <- qnorm(1 - (1 - confidence_level_80) / 2)
z_value_95 <- qnorm(1 - (1 - confidence_level_95) / 2)

confidence_intervals_80 <- data.frame(
  lower_80 = point_predictions - z_value_80 * sigma_hat,
  upper_80 = point_predictions + z_value_80 * sigma_hat
)

confidence_intervals_95 <- data.frame(
  lower_95 = point_predictions - z_value_95 * sigma_hat,
  upper_95 = point_predictions + z_value_95 * sigma_hat
)

# Combine point predictions and confidence intervals into a single dataframe
pred_CI_cart2 <- cbind(point_predictions, confidence_intervals_80, confidence_intervals_95)
```

```{r Q5}
# Get the number of observations used for training
n <- nrow(train_df)

# Make point predictions using the trained model on the test data
point_predictions <- predict(fit.bagcart2, newdata = test_df)

# Get the estimated standard error of the residuals
residuals <- test_df$Average_MSRP_spec - point_predictions
sigma_hat <- sqrt(sum(residuals^2) / (n - length(fit.bagcart2$finalModel$lambda)))

# Calculate the confidence intervals for predictions
confidence_level_80 <- 0.8
confidence_level_95 <- 0.95

z_value_80 <- qnorm(1 - (1 - confidence_level_80) / 2)
z_value_95 <- qnorm(1 - (1 - confidence_level_95) / 2)

confidence_intervals_80 <- data.frame(
  lower_80 = point_predictions - z_value_80 * sigma_hat,
  upper_80 = point_predictions + z_value_80 * sigma_hat
)

confidence_intervals_95 <- data.frame(
  lower_95 = point_predictions - z_value_95 * sigma_hat,
  upper_95 = point_predictions + z_value_95 * sigma_hat
)

# Combine point predictions and confidence intervals into a single dataframe
pred_CI_bagcart2 <- cbind(point_predictions, confidence_intervals_80, confidence_intervals_95)
```

```{r Q5}
# Get the number of observations used for training
n <- nrow(train_df)

# Make point predictions using the trained model on the test data
point_predictions <- predict(fit.rf2, newdata = test_df)

# Get the estimated standard error of the residuals
residuals <- test_df$Average_MSRP_spec - point_predictions
sigma_hat <- sqrt(sum(residuals^2) / (n - length(fit.rf2$finalModel$lambda)))

# Calculate the confidence intervals for predictions
confidence_level_80 <- 0.8
confidence_level_95 <- 0.95

z_value_80 <- qnorm(1 - (1 - confidence_level_80) / 2)
z_value_95 <- qnorm(1 - (1 - confidence_level_95) / 2)

confidence_intervals_80 <- data.frame(
  lower_80 = point_predictions - z_value_80 * sigma_hat,
  upper_80 = point_predictions + z_value_80 * sigma_hat
)

confidence_intervals_95 <- data.frame(
  lower_95 = point_predictions - z_value_95 * sigma_hat,
  upper_95 = point_predictions + z_value_95 * sigma_hat
)

# Combine point predictions and confidence intervals into a single dataframe
pred_CI_rf2 <- cbind(point_predictions, confidence_intervals_80, confidence_intervals_95)
```

```{r Q5}
# Get the number of observations used for training
n <- nrow(train_df)

# Make point predictions using the trained model on the test data
point_predictions <- predict(fit.gbm, newdata = test_df)

# Get the estimated standard error of the residuals
residuals <- test_df$Average_MSRP_spec - point_predictions
sigma_hat <- sqrt(sum(residuals^2) / (n - length(fit.gbm$finalModel$lambda)))

# Calculate the confidence intervals for predictions
confidence_level_80 <- 0.8
confidence_level_95 <- 0.95

z_value_80 <- qnorm(1 - (1 - confidence_level_80) / 2)
z_value_95 <- qnorm(1 - (1 - confidence_level_95) / 2)

confidence_intervals_80 <- data.frame(
  lower_80 = point_predictions - z_value_80 * sigma_hat,
  upper_80 = point_predictions + z_value_80 * sigma_hat
)

confidence_intervals_95 <- data.frame(
  lower_95 = point_predictions - z_value_95 * sigma_hat,
  upper_95 = point_predictions + z_value_95 * sigma_hat
)

# Combine point predictions and confidence intervals into a single dataframe
pred_CI_gbm2 <- cbind(point_predictions, confidence_intervals_80, confidence_intervals_95)
```

```{r Q27}
# data_forecast_ridge <- data.frame(Time = test_df$DATE,
#                    Actual = test_df$Average_MSRP_spec,
#                    Forecasted = pred_CI_ridge)
# 
# data_forecast_lasso2 <- data.frame(Time = test_df$DATE,
#                    Actual = test_df$Average_MSRP_spec,
#                    Forecasted = pred_CI_lasso2)
# 
# data_forecast_enet2 <- data.frame(Time = test_df$DATE,
#                    Actual = test_df$Average_MSRP_spec,
#                    Forecasted = pred_CI_enet2)

data_forecast_cart2 <- data.frame(Time = test_df$DATE,
                   Actual = test_df$Average_MSRP_spec,
                   Forecasted = pred_CI_cart2)

data_forecast_bagcart2 <- data.frame(Time = test_df$DATE,
                   Actual = test_df$Average_MSRP_spec,
                   Forecasted = pred_CI_bagcart2)

data_forecast_rf2 <- data.frame(Time = test_df$DATE,
                   Actual = test_df$Average_MSRP_spec,
                   Forecasted = pred_CI_rf2)

data_forecast_gbm2 <- data.frame(Time = test_df$DATE,
                   Actual = test_df$Average_MSRP_spec,
                   Forecasted = pred_CI_gbm2)
```

```{r Q27}
# Plot the forecasted vs. actual values for Linear model
ggplot(data_forecast_lin, aes(Time)) +
  geom_ribbon(aes(ymin = Forecasted.Lo.95, ymax = Forecasted.Hi.95),
              fill = "#BFD4E9", alpha = 0.5) +
  geom_ribbon(aes(ymin = Forecasted.Lo.80, ymax = Forecasted.Hi.80),
              fill = "#9FC5E8", alpha = 0.5) +
  geom_line(aes(y = Actual, color = "Actual")) +
  geom_line(aes(y = Forecasted.Point.Forecast, color = "Forecasted")) +
  labs(x = "Time", y = "Average_MSRP_spec") +
  scale_color_manual(values = c(Actual = "red", Forecasted = "blue")) +
  theme_minimal() + ggtitle("12 Month Forecast vs. Actual Values for Linear model with Inflation")
```

```{r Q27}
# Accuracy for 12 Month - Linear model
forecast::accuracy(fit_forecast_lin, test_df$Average_MSRP_spec) %>% round(4)
```

```{r Q27}
# Plot the forecasted vs. actual values for ARIMAX model
ggplot(data_forecast_ARIMAX, aes(Time)) +
  geom_ribbon(aes(ymin = Forecasted.Lo.95, ymax = Forecasted.Hi.95),
              fill = "#BFD4E9", alpha = 0.5) +
  geom_ribbon(aes(ymin = Forecasted.Lo.80, ymax = Forecasted.Hi.80),
              fill = "#9FC5E8", alpha = 0.5) +
  geom_line(aes(y = Actual, color = "Actual")) +
  geom_line(aes(y = Forecasted.Point.Forecast, color = "Forecasted")) +
  labs(x = "Time", y = "Average_MSRP_spec") +
  scale_color_manual(values = c(Actual = "red", Forecasted = "blue")) +
  theme_minimal() + ggtitle("12 Month Forecast vs. Actual Values for ARIMAX model with Inflation")
```

```{r Q27}
# Accuracy for 12 Month - ARIMAX model
forecast::accuracy(fit_forecast_ARIMAX, test_df$Average_MSRP_spec) %>% round(4)
```

```{r Q27}
# Plot the forecasted vs. actual values for VAR model
ggplot(data_forecast_VAR, aes(Time)) +
  geom_ribbon(aes(ymin = Forecasted.Lo.95, ymax = Forecasted.Hi.95),
              fill = "#BFD4E9", alpha = 0.5) +
  geom_ribbon(aes(ymin = Forecasted.Lo.80, ymax = Forecasted.Hi.80),
              fill = "#9FC5E8", alpha = 0.5) +
  geom_line(aes(y = Actual, color = "Actual")) +
  geom_line(aes(y = Forecasted.Point.Forecast, color = "Forecasted")) +
  labs(x = "Time", y = "Average_MSRP_spec") +
  scale_color_manual(values = c(Actual = "red", Forecasted = "blue")) +
  theme_minimal() + ggtitle("12 Month Forecast vs. Actual Values for VAR model with Inflation")
```

```{r Q27}
# Accuracy for 12 Month - Linear model
forecast::accuracy(fit_forecast_VAR$forecast$Average_MSRP_spec, test_df$Average_MSRP_spec) %>% round(4)
```

```{r Q27}
# # Plot the forecasted vs. actual values for Ridge Regression model
# ggplot(data_forecast_ridge, aes(Time)) +
#   geom_ribbon(data = data_forecast_ridge, aes(ymin = Forecasted.lower_95, ymax = Forecasted.upper_95),
#               fill = "#BFD4E9", alpha = 0.5) +
#   geom_ribbon(data = data_forecast_ridge, aes(ymin = Forecasted.lower_80, ymax = Forecasted.upper_80),
#               fill = "#9FC5E8", alpha = 0.5) +
#   geom_line(aes(y = Actual, color = "Actual")) +
#   geom_line(aes(y = Forecasted.point_predictions, color = "Forecasted")) +
#   labs(x = "Time", y = "Average MSRP") +
#   scale_color_manual(values = c(Actual = "red", Forecasted = "blue")) +
#   theme_minimal() + ggtitle("12 Month Forecast vs. Actual Values for Ridge Regression model with Inflation")
```

```{r Q27}
# # Accuracy for 12 Month - Lasso Regression model
# forecast::accuracy(pred_CI_ridge$point_predictions, test_df$Average_MSRP_spec) %>% round(4)
```

```{r Q27}
# # Plot the forecasted vs. actual values for Lasso Regression model
# ggplot(data_forecast_lasso2, aes(Time)) +
#   geom_ribbon(data = data_forecast_lasso2, aes(ymin = Forecasted.lower_95, ymax = Forecasted.upper_95),
#               fill = "#BFD4E9", alpha = 0.5) +
#   geom_ribbon(data = data_forecast_lasso2, aes(ymin = Forecasted.lower_80, ymax = Forecasted.upper_80),
#               fill = "#9FC5E8", alpha = 0.5) +
#   geom_line(aes(y = Actual, color = "Actual")) +
#   geom_line(aes(y = Forecasted.point_predictions, color = "Forecasted")) +
#   labs(x = "Time", y = "Average MSRP") +
#   scale_color_manual(values = c(Actual = "red", Forecasted = "blue")) +
#   theme_minimal() + ggtitle("12 Month Forecast vs. Actual Values for Lasso Regression model with Inflation")
```

```{r Q27}
# # Accuracy for 12 Month - Lasso Regression model
# forecast::accuracy(pred_CI_lasso2$point_predictions, test_df$Average_MSRP_spec) %>% round(4)
```

```{r Q27}
# # Plot the forecasted vs. actual values for ElasticNet Regression model
# ggplot(data_forecast_enet2, aes(Time)) +
#   geom_ribbon(data = data_forecast_enet2, aes(ymin = Forecasted.lower_95, ymax = Forecasted.upper_95),
#               fill = "#BFD4E9", alpha = 0.5) +
#   geom_ribbon(data = data_forecast_enet2, aes(ymin = Forecasted.lower_80, ymax = Forecasted.upper_80),
#               fill = "#9FC5E8", alpha = 0.5) +
#   geom_line(aes(y = Actual, color = "Actual")) +
#   geom_line(aes(y = Forecasted.point_predictions, color = "Forecasted")) +
#   labs(x = "Time", y = "Average MSRP") +
#   scale_color_manual(values = c(Actual = "red", Forecasted = "blue")) +
#   theme_minimal() + ggtitle("12 Month Forecast vs. Actual Values for ElasticNet Regression model with Inflation")
```
  
```{r Q27}
# # Accuracy for 12 Month - ElasticNet Regression model
# forecast::accuracy(pred_CI_enet2$point_predictions, test_df$Average_MSRP_spec) %>% round(4)
```

```{r Q27}
# Plot the forecasted vs. actual values for Decision Tree - CART model
ggplot(data_forecast_cart2, aes(Time)) +
  geom_ribbon(data = data_forecast_cart2, aes(ymin = Forecasted.lower_95, ymax = Forecasted.upper_95),
              fill = "#BFD4E9", alpha = 0.5) +
  geom_ribbon(data = data_forecast_cart2, aes(ymin = Forecasted.lower_80, ymax = Forecasted.upper_80),
              fill = "#9FC5E8", alpha = 0.5) +
  geom_line(aes(y = Actual, color = "Actual")) +
  geom_line(aes(y = Forecasted.point_predictions, color = "Forecasted")) +
  labs(x = "Time", y = "Average MSRP") +
  scale_color_manual(values = c(Actual = "red", Forecasted = "blue")) +
  theme_minimal() + ggtitle("12 Month Forecast vs. Actual Values for Decision Tree - CART model with Inflation")
```

```{r Q27}
# Accuracy for 12 Month - Decision Tree - CART model
forecast::accuracy(pred_CI_cart2$point_predictions, test_df$Average_MSRP_spec) %>% round(4)
```

```{r Q27}
# Plot the forecasted vs. actual values for Bagged CART model
ggplot(data_forecast_bagcart2, aes(Time)) +
  geom_ribbon(data = data_forecast_bagcart2, aes(ymin = Forecasted.lower_95, ymax = Forecasted.upper_95),
              fill = "#BFD4E9", alpha = 0.5) +
  geom_ribbon(data = data_forecast_bagcart2, aes(ymin = Forecasted.lower_80, ymax = Forecasted.upper_80),
              fill = "#9FC5E8", alpha = 0.5) +
  geom_line(aes(y = Actual, color = "Actual")) +
  geom_line(aes(y = Forecasted.point_predictions, color = "Forecasted")) +
  labs(x = "Time", y = "Average MSRP") +
  scale_color_manual(values = c(Actual = "red", Forecasted = "blue")) +
  theme_minimal() + ggtitle("12 Month Forecast vs. Actual Values for Bagged CART model with Inflation")
```
  
```{r Q27}
# Accuracy for 12 Month - Bagged CART model
forecast::accuracy(pred_CI_bagcart2$point_predictions, test_df$Average_MSRP_spec) %>% round(4)
```

```{r Q27}
# Plot the forecasted vs. actual values for Random Forest model
ggplot(data_forecast_rf2, aes(Time)) +
  geom_ribbon(data = data_forecast_rf2, aes(ymin = Forecasted.lower_95, ymax = Forecasted.upper_95),
              fill = "#BFD4E9", alpha = 0.5) +
  geom_ribbon(data = data_forecast_rf2, aes(ymin = Forecasted.lower_80, ymax = Forecasted.upper_80),
              fill = "#9FC5E8", alpha = 0.5) +
  geom_line(aes(y = Actual, color = "Actual")) +
  geom_line(aes(y = Forecasted.point_predictions, color = "Forecasted")) +
  labs(x = "Time", y = "Average MSRP") +
  scale_color_manual(values = c(Actual = "red", Forecasted = "blue")) +
  theme_minimal() + ggtitle("12 Month Forecast vs. Actual Values for Random Forest model with Inflation")
```

```{r Q27}
# Accuracy for 12 Month - Random Forest model
forecast::accuracy(pred_CI_rf2$point_predictions, test_df$Average_MSRP_spec) %>% round(4)
```

```{r Q27}
# Plot the forecasted vs. actual values for Gradient Boosting model
ggplot(data_forecast_gbm2, aes(Time)) +
  geom_ribbon(data = data_forecast_gbm2, aes(ymin = Forecasted.lower_95, ymax = Forecasted.upper_95),
              fill = "#BFD4E9", alpha = 0.5) +
  geom_ribbon(data = data_forecast_gbm2, aes(ymin = Forecasted.lower_80, ymax = Forecasted.upper_80),
              fill = "#9FC5E8", alpha = 0.5) +
  geom_line(aes(y = Actual, color = "Actual")) +
  geom_line(aes(y = Forecasted.point_predictions, color = "Forecasted")) +
  labs(x = "Time", y = "Average MSRP") +
  scale_color_manual(values = c(Actual = "red", Forecasted = "blue")) +
  theme_minimal() + ggtitle("12 Month Forecast vs. Actual Values for Gradient Boosting model with Inflation")
```
  
```{r Q27}
# Accuracy for 12 Month - Gradient Boosting model
forecast::accuracy(pred_CI_gbm2$point_predictions, test_df$Average_MSRP_spec) %>% round(4)
```

### Saving model and loading for later use

```{r}
# saveRDS(fit.enet2, "./finalModel_enet2_Regression.rds")
# # fit.enet2 <- readRDS("finalModel_enet2_Regression.rds")
# 
# saveRDS(fit.lasso2, "./finalModel_lasso2_Regression.rds")
# # fit.lasso2 <- readRDS("finalModel_lasso2_Regression.rds")
```